\section{Summary}
We started out trying to get a cluster of Raspberry PIs to run Voldemort. After a few weeks of research, debugging hair pulling we concluded that the PIs were simply too weak and we looked had to look elsewhere for faster hardware. Through our project we have fundamentally changed how Voldemort stores configuration data and utilized ZooKeeper to provide powerful new features. We have created Headmaster to administrate the cluster and allow for new nodes to automatically be included in the cluster. Finally we have created a monitor service using the SigarAPI and the decision taker StatusAnalyzer. Through our experiments we have verified that our implementation does not suffer any noticeable performances loss and successfully migrates partitions to achieve a more balanced cluster. 



% Hva har vi gjort:
% 	- Researcha Voldemort og ZooKeeper
% 		- Papers
% 		- Single node setup
% 	- Prøvd oppsett på PIer
% 		- Issues with memory
% 	- Satt opp ZooKeeper og Voldemort på VMer
% 		- No real issues
% 	- Modified Voldemort to store config in ZooKeeper
% 		- 
% 	- Created functions for Rebalancing and Planning
% 		- 
% 	- Created Headmaster
% 		- System on the outside of Voldemort
% 	- Created Sigar
% 		- Monitor system
% 	- Created Analyzer
% 		- Decission Taker

% 	- Experiments
% 		- Want to verify that we have not broken anything and that stuff works as intended
% 	- Results
% 		- We have done this

\section{Conclusion}

Goals: 

Disposisjon:
	
	- Management is cumbersome. Requires lots of manhours
	- Many system have their own implementations of automatic scaling (Dynamo, Cassandra)

	- We have successfully created such a system for Voldemort. 

	- Automatic management might not have a permanent place in production systems, but they could be used to get a greater understanding for cluster balance

	- 




